---
title: "STAT694_project_House_Pricing"
author: "Haoran Wang"
output:
  html_document:
    df_print: paged
---


## Part 1 Introduction

##Ask a home buyer to describe their dream house, and they probably won’t begin with the height of the basement ceiling or the proximity to an east-west railroad. But this playground competition’s dataset proves that much more influences price negotiations than the number of bedrooms or a white-picket fence. With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, this competition challenges you to predict the final price of each home.
##Here is the Kaggle link: https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data

##Part 2 Loading and Exploring Data

##Loading libraries required and reading the data into R

```{r}
#Loading R packages
library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(xgboost)
library(GGally)
```




```{r}
train <- read.csv("train.csv")
head(train)  ##display first 5 variables and the response variable
```

##Data size and structure

##The train dataset consist of character and integer variables. Most of the character variables are actually (ordinal) factors, but I chose to read them into R as character strings as most of them require cleaning and/or feature engineering first. In total, there are 81 columns/variables, of which the last one is the response variable (SalePrice).

```{r}
dim(train)
```


```{r}
str(train)
```

```{r}
ggplot(data=train[!is.na(train$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="blue", binwidth = 10000) +
        scale_x_continuous(breaks= seq(0, 800000, by=100000), labels = comma) #the sale prices are right skewed
```

##According the plot, we can find the sale prices are right skewed. Most of people tend to but a cheap house so it is resonable.

```{r}
summary(train$SalePrice)
```


##Part 3 Exploring some of the most important variables


##first we should find which numeric variables have a high correlation with the SalePrice.
## In fact, 6 factors that influence a home's value. Factors to consider when pricing a home are: historic sales price, quality of the neighborhood, the market, nearby features and the size, appeal, age and condition of the home. ##The we will dine correlations with SalePrice


```{r}
numericVars <- which(sapply(train, is.numeric)) #index vector numeric variables
numericVarNames <- names(numericVars) #saving names vector for use later on
cat('There are', length(numericVars), 'numeric variables')
```
```{r}
all_numVar <- train[, numericVars]
cor_numVar <- cor(all_numVar, use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")
```

##Overall Quality has the highest correlation with SalePrice among the numeric variables (0.79). It rates the overall material and finish of the house on a scale from 1 (very poor) to 10 (very excellent).

```{r}
pairs(SalePrice ~ OverallQual + GrLivArea + GarageCars + GarageArea + TotalBsmtSF +X1stFlrSF + FullBath+ TotRmsAbvGrd +YearBuilt +YearRemodAdd, data=train)
```

```{r}
ggplot(data=train[!is.na(train$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot(col='blue') + labs(x='Overall Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

##The positive correlation is certainly there indeed, and seems to be a slightly upward curve. Regarding outliers, I do not see any extreme values.



## part 4 Liner regression model 
## according previous output, we will creat our fist regression linear moderl

```{r}
model1 <- lm(SalePrice ~ OverallQual + GrLivArea + GarageCars + GarageArea + TotalBsmtSF +X1stFlrSF + FullBath+ TotRmsAbvGrd +YearBuilt +YearRemodAdd, data=train) 
summary(model1)
```

##According STAT632 knowlwdge, we will remove variables which are not significant by p-value. First we will remove TotRmsAbvGrd due to pvalue is 0.976404 large than 0.05. Then we will get second model


```{r}
model2 <- lm(SalePrice ~ OverallQual + GrLivArea + GarageCars + GarageArea + TotalBsmtSF +X1stFlrSF + FullBath +YearBuilt +YearRemodAdd, data=train) 
summary(model2)
```

##next we will remove GarageArea due to pvalue is 0.146838 large than 0.05. Then we will get 3ed model

```{r}
model3 <- lm(SalePrice ~ OverallQual + GrLivArea + GarageCars  + TotalBsmtSF +X1stFlrSF + FullBath +YearBuilt +YearRemodAdd, data=train) 
summary(model3)
```

##model 3 looks better, all variables are significant 


```{r}
par(mfrow=c(1,2), cex=0.6) 
plot(model3, 1:2)  #check assumption
```


```{r}
summary(model3)$adj.r.squared
```

```{r}
q1 <- ggplot(train, aes(OverallQual,SalePrice)) + 
  geom_point()
q2 <- ggplot(train, aes( GrLivArea,SalePrice)) + 
  geom_point()
q3 <- ggplot(train, aes( GarageCars,SalePrice)) + 
  geom_point()
q4 <- ggplot(train, aes( GarageArea,SalePrice)) + 
  geom_point()
q5 <- ggplot(train, aes( YearRemodAdd,SalePrice)) + 
  geom_point() 
  
q6 <- ggplot(train, aes( X1stFlrSF,SalePrice)) + 
  geom_point()
q7 <- ggplot(train, aes( FullBath,SalePrice)) + 
  geom_point()
q8 <- ggplot(train, aes( TotRmsAbvGrd,SalePrice)) + 
  geom_point()


layout <- matrix(c(1,2,3,4,5,9,6,7,8),3,3,byrow=TRUE)
multiplot(q1, q2, q3, q4, q5, q6, q7,q8, layout=layout)

```

##Overall Quality is very important, and also more granular than the other variables. 
##Most of house has 1 to 2 bath room, it is different out recently market, may be because their bulid year is old. 

##now we are analyze two main factors : OverallQual ,GrLivArea.

```{r}
train2 <- select(train,  SalePrice, OverallQual ,GrLivArea)
head(train2) #new data set 
```

```{r}
ggplot(train2, aes(SalePrice, GrLivArea)) + 
  geom_point()
```

```{r}
ggplot(train2, aes(SalePrice, OverallQual)) + 
  geom_point()
```


```{r}
ggplot(train2, aes(SalePrice, GrLivArea, colour = OverallQual)) + 
  geom_point()
```


##let we use "filter" function, and then find all obs which OverallQual == 5, then we will analye the SalePrice and GrLivArea. We also can use other levels OverallQual


```{r}
train3 <- filter(train2,   OverallQual == 8)
ggplot(train3, aes(GrLivArea,SalePrice)) + 
  geom_point()
```

```{r}
train4 <- filter(train2,   OverallQual == 7)
ggplot(train4, aes(GrLivArea,SalePrice)) + 
  geom_point()
```



```{r}
train5 <- filter(train2,   OverallQual == 5)
ggplot(train5, aes( GrLivArea,SalePrice)) + 
  geom_point()
```

##According output plots, we can simply say in higl level OverallQual property, Salesprice has stronger liner relationship with Grlivearea. 

## Summary: In this project, I tried to find the most siginificant factors and creat a regression model. Of course, a house pricing can not only calculate by a few factors. There are so may factors may influence market price. But I reviewed linear regression and data visualization via this projet, it is a good experience for me. 
